---
title: "Octopus survey results"
author: "Andrew Mertens"
date: "2025-07-12"
output:
  html_document: default
---


```{r setupALSE}
knitr::opts_chunk$set(echo = FALSE)

rm(list=ls())
library(tidyverse)
library(here)
library(table1)
library(DT)
library(flextable)

d <- readRDS(here("data", "cleaned_octopus_farming_data.RDS"))
colnames(d)
dput(colnames(d))
```


## Explore demographics by country

```{r}
table1(~ Gender + Age + dAGE + URBANRURAL + income_cat | COUNTRY, data=d, transpose=TRUE)
```




## Explore opinions by country

```{r}
table1(~ Q1_awareness + Q2_support + Q4_welfare_knowledge + Q4_environmental_knowledge + Q4_economic_knowledge + 
         Q4_sentience_knowledge + Q4_public_health_knowledge + Q5_consumption_frequency + Q6_impact_on_consumption + 
         Q7_ban_farming_support + Q7_ban_sale_support + Q7_ban_subsidies_support + Q9_vote_candidate_support| COUNTRY, data=d, transpose=FALSE)
```


## Explore support by country


```{r}


# [1] Neutral / No opinion Strongly oppose      Slightly support    
# [4] Moderately support   Don't know           Moderately oppose   
# [7] Strongly support     Slightly oppose 

d <- d %>%
  mutate(support = case_when(
    Q2_support == "Strongly support" ~ "Support",
    Q2_support == "Moderately support" ~ "Support",
    Q2_support == "Slightly support" ~ "Support",
    Q2_support == "Neutral / No opinion" ~ "Neutral/Don't know",
    Q2_support == "Don't know" ~ "Neutral/Don't know",
    Q2_support == "Slightly oppose" ~ "Oppose",
    Q2_support == "Moderately oppose" ~ "Oppose",
    Q2_support == "Strongly oppose" ~ "Oppose"),
    
    ban_farming_support = case_when(
    Q7_ban_farming_support == "Strongly support" ~ "Support",
    Q7_ban_farming_support == "Moderately support" ~ "Support",
    Q7_ban_farming_support == "Slightly support" ~ "Support",
    Q7_ban_farming_support == "Neutral / No opinion" ~ "Neutral/Don't know",
    Q7_ban_farming_support == "Don't know" ~ "Neutral/Don't know",
    Q7_ban_farming_support == "Slightly oppose" ~ "Oppose",
    Q7_ban_farming_support == "Moderately oppose" ~ "Oppose",
    Q7_ban_farming_support == "Strongly oppose" ~ "Oppose"),
    
    ban_sale_support = case_when(
    Q7_ban_sale_support == "Strongly support" ~ "Support",
    Q7_ban_sale_support == "Moderately support" ~ "Support",
    Q7_ban_sale_support == "Slightly support" ~ "Support",
    Q7_ban_sale_support == "Neutral / No opinion" ~ "Neutral/Don't know",
    Q7_ban_sale_support == "Don't know" ~ "Neutral/Don't know",
    Q7_ban_sale_support == "Slightly oppose" ~ "Oppose",
    Q7_ban_sale_support == "Moderately oppose" ~ "Oppose",
    Q7_ban_sale_support == "Strongly oppose" ~ "Oppose"),
    
    ban_subsidies_support = case_when(
    Q7_ban_sale_support == "Strongly support" ~ "Support",
    Q7_ban_sale_support == "Moderately support" ~ "Support",
    Q7_ban_sale_support == "Slightly support" ~ "Support",
    Q7_ban_sale_support == "Neutral / No opinion" ~ "Neutral/Don't know",
    Q7_ban_sale_support == "Don't know" ~ "Neutral/Don't know",
    Q7_ban_sale_support == "Slightly oppose" ~ "Oppose",
    Q7_ban_sale_support == "Moderately oppose" ~ "Oppose",
    Q7_ban_sale_support == "Strongly oppose" ~ "Oppose")
  )



cat("Support farming octopuses(pre-message):\n")
tab=round(prop.table(table(d$support))*100,1)
knitr::kable((tab))
tab=round(prop.table(table(d$COUNTRY, d$support),1)*100,1) 
knitr::kable(as.data.frame.matrix(tab) %>% arrange(Support))

cat("Support for ban on farming octopuses (post-message): \n")
tab=round(prop.table(table(d$ban_farming_support))*100,1)
knitr::kable((tab) )
tab=round(prop.table(table(d$COUNTRY, d$ban_farming_support),1)*100,1)
knitr::kable(as.data.frame.matrix(tab) %>% arrange(Support))

cat("Support for ban on sale of octopuses (post-message): \n")
tab=round(prop.table(table(d$ban_sale_support))*100,1)
knitr::kable((tab))
tab=round(prop.table(table(d$COUNTRY, d$ban_sale_support),1)*100,1)
knitr::kable(as.data.frame.matrix(tab) %>% arrange(Support))

cat("Support for ban on subsidies (post-message): \n")
tab=round(prop.table(table(d$ban_subsidies_support))*100,1)
knitr::kable((tab))
tab=round(prop.table(table(d$COUNTRY, d$ban_subsidies_support),1)*100,1)
knitr::kable(as.data.frame.matrix(tab) %>% arrange(Support))

```


## Explore demographics by message

```{r}
table1(~ Gender + Age + dAGE + URBANRURAL + income_cat | message, data=d, transpose=FALSE)
```

## Explore opinions by message

```{r}
table1(~ Q1_awareness + Q2_support + Q4_welfare_knowledge + Q4_environmental_knowledge + Q4_economic_knowledge + 
         Q4_sentience_knowledge + Q4_public_health_knowledge + Q5_consumption_frequency + Q6_impact_on_consumption + 
         Q7_ban_farming_support + Q7_ban_sale_support + Q7_ban_subsidies_support + Q9_vote_candidate_support| message, data=d, transpose=FALSE)
```

## Explore opinions by message - coded as continious

```{r}

d[d=="Don't know"] <- NA

d <- d %>%
  mutate(Q1_awareness=as.numeric(Q1_awareness),
         Q2_support=as.numeric(Q2_support),
         Q4_welfare_knowledge=as.numeric(Q4_welfare_knowledge),
         Q4_environmental_knowledge=as.numeric(Q4_environmental_knowledge),
         Q4_economic_knowledge=as.numeric(Q4_economic_knowledge),
         Q4_sentience_knowledge=as.numeric(Q4_sentience_knowledge),
         Q4_public_health_knowledge=as.numeric(Q4_public_health_knowledge),
         Q5_consumption_frequency=as.numeric(Q5_consumption_frequency),
         Q6_impact_on_consumption=as.numeric(Q6_impact_on_consumption),
         Q7_ban_farming_support=as.numeric(Q7_ban_farming_support),
         Q7_ban_sale_support=as.numeric(Q7_ban_sale_support),
         Q7_ban_subsidies_support=as.numeric(Q7_ban_subsidies_support),
         Q9_vote_candidate_support=as.numeric(Q9_vote_candidate_support))
table1(~ Q1_awareness + Q2_support + Q4_welfare_knowledge + Q4_environmental_knowledge + Q4_economic_knowledge + 
         Q4_sentience_knowledge + Q4_public_health_knowledge + Q5_consumption_frequency + Q6_impact_on_consumption + 
         Q7_ban_farming_support + Q7_ban_sale_support + Q7_ban_subsidies_support + Q9_vote_candidate_support| message, data=d, transpose=FALSE)
```



## Primary results- effects of message conditions adjusting for covariates


```{r}

res_ban_farming_adj = readRDS(here("results/res_farming_unadj.rds"))
res_ban_sale_adj = readRDS(here("results/res_sale_unadj.rds"))
res_ban_subsidies_adj = readRDS(here("results/res_subsidies_unadj.rds"))

res_ban_farming_adj$Y = gsub("tr","",res_ban_farming_adj$Y)
res_ban_sale_adj$Y = gsub("tr","",res_ban_sale_adj$Y)
res_ban_subsidies_adj$Y = gsub("tr","",res_ban_subsidies_adj$Y)

ggplot(res_ban_farming_adj, aes(x=Y, y=`Coef.`)) + geom_point() +
  geom_errorbar(aes(ymin=`2.5%`, ymax=`97.5%`), width=0.2) +
  geom_hline(yintercept=0, linetype="dashed", color="red") +
  labs(title="Effect of Message on Support\nfor Ban on Farming Octopuses",
       x="Message Condition",
       y="Estimated Effect (95% CI)") +
  coord_flip() +
  theme_minimal() 

#datatable(res_ban_farming_adj, options = list(pageLength = 5, autoWidth = TRUE))
knitr::kable(res_ban_farming_adj, digits=2)


ggplot(res_ban_sale_adj, aes(x=Y, y=`Coef.`)) + geom_point() +
  geom_errorbar(aes(ymin=`2.5%`, ymax=`97.5%`), width=0.2) +
  geom_hline(yintercept=0, linetype="dashed", color="red") +
  labs(title="Effect of Message on Support\nfor Ban on Sale of Octopuses",
       x="Message Condition",
       y="Estimated Effect (95% CI)") +
  coord_flip() +
  theme_minimal() 
#datatable(res_ban_sale_adj, options = list(pageLength = 5, autoWidth = TRUE))
knitr::kable(res_ban_sale_adj, digits=2)

ggplot(res_ban_subsidies_adj, aes(x=Y, y=`Coef.`)) + geom_point() +
  geom_errorbar(aes(ymin=`2.5%`, ymax=`97.5%`), width=0.2) +
  geom_hline(yintercept=0, linetype="dashed", color="red") +
  labs(title="Effect of Message on Support\nfor Ban on Subsidies",
       x="Message Condition",
       y="Estimated Effect (95% CI)") +
  coord_flip() +
  theme_minimal() 
#datatable(res_ban_subsidies_adj, options = list(pageLength = 5, autoWidth = TRUE))
knitr::kable(res_ban_subsidies_adj, digits=2)

```


## Secondary results- secondary message contrasts

```{r}
res_ban_farming_adj = readRDS(here("results/res_farming_secondary.rds"))
res_ban_sale_adj = readRDS(here("results/res_sale_secondary.rds"))
res_ban_subsidies_adj = readRDS(here("results/res_subsidies_secondary.rds"))

res_ban_farming_adj$contrast = gsub(" \\(ref\\)","",res_ban_farming_adj$contrast)
res_ban_sale_adj$contrast = gsub(" \\(ref\\)","",res_ban_sale_adj$contrast)
res_ban_subsidies_adj$contrast = gsub(" \\(ref\\)","",res_ban_subsidies_adj$contrast)


ggplot(res_ban_farming_adj, aes(x=contrast, y=`Coef.`)) + geom_point() +
  geom_errorbar(aes(ymin=`2.5%`, ymax=`97.5%`), width=0.2) +
  geom_hline(yintercept=0, linetype="dashed", color="red") +
  labs(title="Effect of Message on Support\nfor Ban on Farming Octopuses",
       x="Message Condition Contrast",
       y="Estimated Effect (95% CI)") +
  coord_flip() +
  theme_minimal() 

ggplot(res_ban_sale_adj, aes(x=contrast, y=`Coef.`)) + geom_point() +
  geom_errorbar(aes(ymin=`2.5%`, ymax=`97.5%`), width=0.2) +
  geom_hline(yintercept=0, linetype="dashed", color="red") +
  labs(title="Effect of Message on Support\nfor Ban on Sale of Octopuses",
       x="Message Condition Contrast",
       y="Estimated Effect (95% CI)") +
  coord_flip() +
  theme_minimal() 

ggplot(res_ban_subsidies_adj, aes(x=contrast, y=`Coef.`)) + geom_point() +
  geom_errorbar(aes(ymin=`2.5%`, ymax=`97.5%`), width=0.2) +
  geom_hline(yintercept=0, linetype="dashed", color="red") +
  labs(title="Effect of Message on Support\nfor Ban on Subsidies",
       x="Message Condition Contrast",
       y="Estimated Effect (95% CI)") +
  coord_flip() +
  theme_minimal() 
```


## Secondary analysis- country-specific message effects

```{r}

res_ban_farming_subgroup = readRDS(here("results/res_farming_subgroup.rds")) %>% rename(Message=tr)
res_ban_sale_subgroup = readRDS(here("results/res_sale_subgroup.rds")) %>% rename(Message=tr)
res_ban_subsidies_subgroup = readRDS(here("results/res_subsidies_subgroup.rds")) %>% rename(Message=tr)

ggplot(res_ban_farming_subgroup, aes(x=Country, y=est, group=Message, color=Message)) + geom_point(position = position_dodge(0.5)) +
  geom_errorbar(aes(ymin=est.lb, ymax=est.ub), width=0.2, position = position_dodge(0.5)) +
  geom_hline(yintercept=0, linetype="dashed", color="red") +
  labs(title="Effect of Message on Support\nfor Ban on Farming Octopuses",
       x="Country",
       y="Estimated Effect (95% CI)") +
  coord_flip() +
  theme_minimal() 

ggplot(res_ban_sale_subgroup, aes(x=Country, y=est, group=Message, color=Message)) + 
  geom_point(position = position_dodge(0.5)) +
  geom_errorbar(aes(ymin=est.lb, ymax=est.ub), width=0.2, position = position_dodge(0.5)) +
  geom_hline(yintercept=0, linetype="dashed", color="red") +
  labs(title="Effect of Message on Support\nfor Ban on Sale of Octopuses",
       x="Country",
       y="Estimated Effect (95% CI)") +
  coord_flip() +
  theme_minimal() 

ggplot(res_ban_subsidies_subgroup, aes(x=Country, y=est, group=Message, color=Message)) + 
    geom_point(position = position_dodge(0.5)) +
  geom_errorbar(aes(ymin=est.lb, ymax=est.ub), width=0.2, position = position_dodge(0.5)) +
  geom_hline(yintercept=0, linetype="dashed", color="red") +
  labs(title="Effect of Message on Support\nfor Ban on Subsidies",
       x="Country",
       y="Estimated Effect (95% CI)") +
  coord_flip() +
  theme_minimal() 



```


## Variable importance analysis- multivariate associations

```{r}
plotdf_Y1 = readRDS(here("results/res_farm_vim_glm.rds"))
plotdf_Y2 = readRDS(here("results/res_sale_vim_glm.rds"))
plotdf_Y3 = readRDS(here("results/res_subsidies_vim_glm.rds"))


clean_df <- function(d){
  d$X = gsub("_"," ",d$X) 
  d$X = gsub("  "," ",d$X) 
  d$X = gsub("Q1 ","",d$X) 
  d$X = gsub("Q2 ","",d$X) 
  d$X = gsub("Q4 ","",d$X) 
  d$X = gsub("Q6 ","",d$X) 
  d$X = gsub("Q5 ","",d$X) 
  d$X = gsub("COUNTRY","Country",d$X) 
  d$X = gsub("dAGE","Age",d$X) 
  d$X = gsub("dAGE","Age",d$X) 
  d$X = gsub("URBANRURAL: ","Location: ",d$X) 
  d$X = gsub("income_cat","Income",d$X) 
  d$X =str_to_title(d$X)
  return(d)
}

plotdf_Y1 = clean_df(plotdf_Y1)
plotdf_Y2 = clean_df(plotdf_Y2)
plotdf_Y3 = clean_df(plotdf_Y3)

ggplot(plotdf_Y1, aes(x=X, y=est )) + geom_point() +
  geom_errorbar(aes(ymin=ci.lb, ymax=ci.ub), width=0.5) +
  geom_hline(yintercept=0, linetype="dashed", color="red") +
  labs(title="Support for octopus farm ban",
    x="Covariate",
    y="Estimated Effect (95% CI)") +
  coord_flip() +theme_minimal() 

ggplot(plotdf_Y2, aes(x=X, y=est )) + geom_point() +
  geom_errorbar(aes(ymin=ci.lb, ymax=ci.ub), width=0.5) +
  geom_hline(yintercept=0, linetype="dashed", color="red") +
  labs(title="Support for octopus sale ban",
    x="Covariate",
    y="Estimated Effect (95% CI)") +
  coord_flip() +theme_minimal() 

ggplot(plotdf_Y3, aes(x=X, y=est )) + geom_point() +
  geom_errorbar(aes(ymin=ci.lb, ymax=ci.ub), width=0.5) +
  geom_hline(yintercept=0, linetype="dashed", color="red") +
  labs(title="Support for ban on subsidies",
    x="Covariate",
    y="Estimated Effect (95% CI)") +
  coord_flip() +theme_minimal() 

```

## Sensitivity results- Variable importance analysis- bivariate associations

```{r}
plotdf_Y1 = readRDS(here("results/res_farm_vim_bivariate.rds"))
plotdf_Y2 = readRDS(here("results/res_sale_vim_bivariate.rds"))
plotdf_Y3 = readRDS(here("results/res_subsidies_vim_bivariate.rds"))



plotdf_Y1 = clean_df(plotdf_Y1)
plotdf_Y2 = clean_df(plotdf_Y2)
plotdf_Y3 = clean_df(plotdf_Y3)

ggplot(plotdf_Y1, aes(x=X, y=est )) + geom_point() +
  geom_errorbar(aes(ymin=ci.lb, ymax=ci.ub), width=0.5) +
  geom_hline(yintercept=0, linetype="dashed", color="red") +
  labs(title="Support for octopus farm ban",
    x="Covariate",
    y="Estimated Effect (95% CI)") +
  coord_flip() +theme_minimal() 

ggplot(plotdf_Y2, aes(x=X, y=est )) + geom_point() +
  geom_errorbar(aes(ymin=ci.lb, ymax=ci.ub), width=0.5) +
  geom_hline(yintercept=0, linetype="dashed", color="red") +
  labs(title="Support for octopus sale ban",
    x="Covariate",
    y="Estimated Effect (95% CI)") +
  coord_flip() +theme_minimal() 

ggplot(plotdf_Y3, aes(x=X, y=est )) + geom_point() +
  geom_errorbar(aes(ymin=ci.lb, ymax=ci.ub), width=0.5) +
  geom_hline(yintercept=0, linetype="dashed", color="red") +
  labs(title="Support for ban on subsidies",
    x="Covariate",
    y="Estimated Effect (95% CI)") +
  coord_flip() +theme_minimal() 

```

## Sensitivity results- unadjusted effects of message conditions 

```{r}

res_ban_farming_unadj = readRDS(here("results/res_farming_unadj.rds"))
res_ban_sale_unadj = readRDS(here("results/res_sale_unadj.rds"))
res_ban_subsidies_unadj = readRDS(here("results/res_subsidies_unadj.rds"))

res_ban_farming_unadj$Y = gsub("tr","",res_ban_farming_unadj$Y)
res_ban_sale_unadj$Y = gsub("tr","",res_ban_sale_unadj$Y)
res_ban_subsidies_unadj$Y = gsub("tr","",res_ban_subsidies_unadj$Y)


ggplot(res_ban_farming_unadj, aes(x=Y, y=`Coef.`)) + geom_point() +
  geom_errorbar(aes(ymin=`2.5%`, ymax=`97.5%`), width=0.2) +
  geom_hline(yintercept=0, linetype="dashed", color="red") +
  labs(title="Effect of Message on Support\nfor Ban on Farming Octopuses",
       x="Message Condition",
       y="Estimated Effect (95% CI)") +
  coord_flip() +
  theme_minimal() 

ggplot(res_ban_sale_unadj, aes(x=Y, y=`Coef.`)) + geom_point() +
  geom_errorbar(aes(ymin=`2.5%`, ymax=`97.5%`), width=0.2) +
  geom_hline(yintercept=0, linetype="dashed", color="red") +
  labs(title="Effect of Message on Support\nfor Ban on Sale of Octopuses",
       x="Message Condition",
       y="Estimated Effect (95% CI)") +
  coord_flip() +
  theme_minimal() 

ggplot(res_ban_subsidies_unadj, aes(x=Y, y=`Coef.`)) + geom_point() +
  geom_errorbar(aes(ymin=`2.5%`, ymax=`97.5%`), width=0.2) +
  geom_hline(yintercept=0, linetype="dashed", color="red") +
  labs(title="Effect of Message on Support\nfor Ban on Subsidies",
       x="Message Condition",
       y="Estimated Effect (95% CI)") +
  coord_flip() +
  theme_minimal() 

```


## Sensitivity results- comparing age and sex distributions by country to EU census

-comparing against Eurostat 2024 data. Below is the summary of the ratio of survey proportions to EU proptions, with the IQR being a small range of  0.9716 to  1.0667, showing most strata are very representative, as seen in the heatmap. The cells that are off are the older ages generally (which probably reflects the fact that older people are less likely to use social media, and thus less likely to be surveyed). Note UK is not in the EU data anymore because of Brexit.


```{r, echo=FALSE}

res_ban_farming_unadj = readRDS(here("results/res_farming_unadj.rds"))
merged = read.csv(here("results/survey_vs_population_representativeness.csv"))


print(summary(merged$representation_ratio))

# Step 1: Use the merged data (after join and ratio calculation)
# If you've already run up to `merged <- ...`, use this:

heat_data <- merged %>%
  mutate(
    label = paste0(country, " (", sex, ")"),
    label = fct_reorder(label, as.numeric(factor(country))*10 + ifelse(sex == "F", 1, 0))
  ) %>%
  select(label, age_band, representation_ratio)

# Step 2: Plot
ggplot(heat_data, aes(x = age_band, y = label, fill = representation_ratio)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "firebrick", mid = "white", high = "steelblue",
                       midpoint = 1, limits = c(0.5, 1.5), oob = scales::squish,
                       name = "Survey / Census") +
  labs(
    title = "Survey Representativeness vs Census Weights (by Sex × Age × Country)",
    x = "Age Band", y = "Country (Sex)"
  ) +
  theme_minimal(base_size = 12) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    panel.grid = element_blank()
  )

```

