---
title: "Octopus survey results"
author: "Andrew Mertens"
date: "2025-07-12"
output:
  html_document: default
---


```{r setupALSE}
knitr::opts_chunk$set(echo = FALSE)

rm(list=ls())
library(tidyverse)
library(here)
library(table1)
library(DT)
library(flextable)

d <- readRDS(here("data", "cleaned_octopus_farming_data.RDS"))
colnames(d)
dput(colnames(d))
```


## Explore demographics by country

```{r}
table1(~ Gender + Age + dAGE + URBANRURAL + income_cat | COUNTRY, data=d, transpose=TRUE)
```




## Explore opinions by country

```{r}
table1(~ Q1_awareness + Q2_support + Q4_welfare_knowledge + Q4_environmental_knowledge + Q4_economic_knowledge + 
         Q4_sentience_knowledge + Q4_public_health_knowledge + Q5_consumption_frequency + Q6_impact_on_consumption + 
         Q7_ban_farming_support + Q7_ban_sale_support + Q7_ban_subsidies_support + Q9_vote_candidate_support| COUNTRY, data=d, transpose=FALSE)
```


## Explore support by country


```{r}


# [1] Neutral / No opinion Strongly oppose      Slightly support    
# [4] Moderately support   Don't know           Moderately oppose   
# [7] Strongly support     Slightly oppose 

d <- d %>%
  mutate(support = case_when(
    Q2_support == "Strongly support" ~ "Support",
    Q2_support == "Moderately support" ~ "Support",
    Q2_support == "Slightly support" ~ "Support",
    Q2_support == "Neutral / No opinion" ~ "Neutral/Don't know",
    Q2_support == "Don't know" ~ "Neutral/Don't know",
    Q2_support == "Slightly oppose" ~ "Oppose",
    Q2_support == "Moderately oppose" ~ "Oppose",
    Q2_support == "Strongly oppose" ~ "Oppose"),
    
    ban_farming_support = case_when(
    Q7_ban_farming_support == "Strongly support" ~ "Support",
    Q7_ban_farming_support == "Moderately support" ~ "Support",
    Q7_ban_farming_support == "Slightly support" ~ "Support",
    Q7_ban_farming_support == "Neutral / No opinion" ~ "Neutral/Don't know",
    Q7_ban_farming_support == "Don't know" ~ "Neutral/Don't know",
    Q7_ban_farming_support == "Slightly oppose" ~ "Oppose",
    Q7_ban_farming_support == "Moderately oppose" ~ "Oppose",
    Q7_ban_farming_support == "Strongly oppose" ~ "Oppose"),
    
    ban_sale_support = case_when(
    Q7_ban_sale_support == "Strongly support" ~ "Support",
    Q7_ban_sale_support == "Moderately support" ~ "Support",
    Q7_ban_sale_support == "Slightly support" ~ "Support",
    Q7_ban_sale_support == "Neutral / No opinion" ~ "Neutral/Don't know",
    Q7_ban_sale_support == "Don't know" ~ "Neutral/Don't know",
    Q7_ban_sale_support == "Slightly oppose" ~ "Oppose",
    Q7_ban_sale_support == "Moderately oppose" ~ "Oppose",
    Q7_ban_sale_support == "Strongly oppose" ~ "Oppose"),
    
    ban_subsidies_support = case_when(
    Q7_ban_sale_support == "Strongly support" ~ "Support",
    Q7_ban_sale_support == "Moderately support" ~ "Support",
    Q7_ban_sale_support == "Slightly support" ~ "Support",
    Q7_ban_sale_support == "Neutral / No opinion" ~ "Neutral/Don't know",
    Q7_ban_sale_support == "Don't know" ~ "Neutral/Don't know",
    Q7_ban_sale_support == "Slightly oppose" ~ "Oppose",
    Q7_ban_sale_support == "Moderately oppose" ~ "Oppose",
    Q7_ban_sale_support == "Strongly oppose" ~ "Oppose")
  )



cat("Support farming octopuses(pre-message):\n")
tab=round(prop.table(table(d$support))*100,1)
knitr::kable((tab))
tab=round(prop.table(table(d$COUNTRY, d$support),1)*100,1) 
knitr::kable(as.data.frame.matrix(tab) %>% arrange(Support))

cat("Support for ban on farming octopuses (post-message): \n")
tab=round(prop.table(table(d$ban_farming_support))*100,1)
knitr::kable((tab) )
tab=round(prop.table(table(d$COUNTRY, d$ban_farming_support),1)*100,1)
knitr::kable(as.data.frame.matrix(tab) %>% arrange(Support))

cat("Support for ban on sale of octopuses (post-message): \n")
tab=round(prop.table(table(d$ban_sale_support))*100,1)
knitr::kable((tab))
tab=round(prop.table(table(d$COUNTRY, d$ban_sale_support),1)*100,1)
knitr::kable(as.data.frame.matrix(tab) %>% arrange(Support))

cat("Support for ban on subsidies (post-message): \n")
tab=round(prop.table(table(d$ban_subsidies_support))*100,1)
knitr::kable((tab))
tab=round(prop.table(table(d$COUNTRY, d$ban_subsidies_support),1)*100,1)
knitr::kable(as.data.frame.matrix(tab) %>% arrange(Support))

```


## Explore demographics by message

```{r}
table1(~ Gender + Age + dAGE + URBANRURAL + income_cat | message, data=d, transpose=FALSE)
```

## Explore opinions by message

```{r}
table1(~ Q1_awareness + Q2_support + Q4_welfare_knowledge + Q4_environmental_knowledge + Q4_economic_knowledge + 
         Q4_sentience_knowledge + Q4_public_health_knowledge + Q5_consumption_frequency + Q6_impact_on_consumption + 
         Q7_ban_farming_support + Q7_ban_sale_support + Q7_ban_subsidies_support + Q9_vote_candidate_support| message, data=d, transpose=FALSE)
```

## Explore opinions by message - coded as continious

```{r}

d[d=="Don't know"] <- NA

d <- d %>%
  mutate(Q1_awareness=as.numeric(Q1_awareness),
         Q2_support=as.numeric(Q2_support),
         Q4_welfare_knowledge=as.numeric(Q4_welfare_knowledge),
         Q4_environmental_knowledge=as.numeric(Q4_environmental_knowledge),
         Q4_economic_knowledge=as.numeric(Q4_economic_knowledge),
         Q4_sentience_knowledge=as.numeric(Q4_sentience_knowledge),
         Q4_public_health_knowledge=as.numeric(Q4_public_health_knowledge),
         Q5_consumption_frequency=as.numeric(Q5_consumption_frequency),
         Q6_impact_on_consumption=as.numeric(Q6_impact_on_consumption),
         Q7_ban_farming_support=as.numeric(Q7_ban_farming_support),
         Q7_ban_sale_support=as.numeric(Q7_ban_sale_support),
         Q7_ban_subsidies_support=as.numeric(Q7_ban_subsidies_support),
         Q9_vote_candidate_support=as.numeric(Q9_vote_candidate_support))
table1(~ Q1_awareness + Q2_support + Q4_welfare_knowledge + Q4_environmental_knowledge + Q4_economic_knowledge + 
         Q4_sentience_knowledge + Q4_public_health_knowledge + Q5_consumption_frequency + Q6_impact_on_consumption + 
         Q7_ban_farming_support + Q7_ban_sale_support + Q7_ban_subsidies_support + Q9_vote_candidate_support| message, data=d, transpose=FALSE)
```



## Primary results- effects of message conditions adjusting for covariates


```{r}

res_ban_farming_adj = readRDS(here("results/res_farming_unadj.rds"))
res_ban_sale_adj = readRDS(here("results/res_sale_unadj.rds"))
res_ban_subsidies_adj = readRDS(here("results/res_subsidies_unadj.rds"))

ggplot(res_ban_farming_adj, aes(x=Y, y=`Coef.`)) + geom_point() +
  geom_errorbar(aes(ymin=`2.5%`, ymax=`97.5%`), width=0.2) +
  geom_hline(yintercept=0, linetype="dashed", color="red") +
  labs(title="Effect of Message on Support\nfor Ban on Farming Octopuses",
       x="Message Condition",
       y="Estimated Effect (95% CI)") +
  coord_flip() +
  theme_minimal() 

#datatable(res_ban_farming_adj, options = list(pageLength = 5, autoWidth = TRUE))
knitr::kable(res_ban_farming_adj, digits=2)


ggplot(res_ban_sale_adj, aes(x=Y, y=`Coef.`)) + geom_point() +
  geom_errorbar(aes(ymin=`2.5%`, ymax=`97.5%`), width=0.2) +
  geom_hline(yintercept=0, linetype="dashed", color="red") +
  labs(title="Effect of Message on Support\nfor Ban on Sale of Octopuses",
       x="Message Condition",
       y="Estimated Effect (95% CI)") +
  coord_flip() +
  theme_minimal() 
#datatable(res_ban_sale_adj, options = list(pageLength = 5, autoWidth = TRUE))
knitr::kable(res_ban_sale_adj, digits=2)

ggplot(res_ban_subsidies_adj, aes(x=Y, y=`Coef.`)) + geom_point() +
  geom_errorbar(aes(ymin=`2.5%`, ymax=`97.5%`), width=0.2) +
  geom_hline(yintercept=0, linetype="dashed", color="red") +
  labs(title="Effect of Message on Support\nfor Ban on Subsidies",
       x="Message Condition",
       y="Estimated Effect (95% CI)") +
  coord_flip() +
  theme_minimal() 
#datatable(res_ban_subsidies_adj, options = list(pageLength = 5, autoWidth = TRUE))
knitr::kable(res_ban_subsidies_adj, digits=2)

```


## Secondary results- secondary message contrasts

```{r}
res_ban_farming_adj = readRDS(here("results/res_farming_secondary.rds"))
res_ban_sale_adj = readRDS(here("results/res_sale_secondary.rds"))
res_ban_subsidies_adj = readRDS(here("results/res_subsidies_secondary.rds"))

ggplot(res_ban_farming_adj, aes(x=contrast, y=`Coef.`)) + geom_point() +
  geom_errorbar(aes(ymin=`2.5%`, ymax=`97.5%`), width=0.2) +
  geom_hline(yintercept=0, linetype="dashed", color="red") +
  labs(title="Effect of Message on Support\nfor Ban on Farming Octopuses",
       x="Message Condition",
       y="Estimated Effect (95% CI)") +
  coord_flip() +
  theme_minimal() 

ggplot(res_ban_sale_adj, aes(x=contrast, y=`Coef.`)) + geom_point() +
  geom_errorbar(aes(ymin=`2.5%`, ymax=`97.5%`), width=0.2) +
  geom_hline(yintercept=0, linetype="dashed", color="red") +
  labs(title="Effect of Message on Support\nfor Ban on Sale of Octopuses",
       x="Message Condition",
       y="Estimated Effect (95% CI)") +
  coord_flip() +
  theme_minimal() 

ggplot(res_ban_subsidies_adj, aes(x=contrast, y=`Coef.`)) + geom_point() +
  geom_errorbar(aes(ymin=`2.5%`, ymax=`97.5%`), width=0.2) +
  geom_hline(yintercept=0, linetype="dashed", color="red") +
  labs(title="Effect of Message on Support\nfor Ban on Subsidies",
       x="Message Condition",
       y="Estimated Effect (95% CI)") +
  coord_flip() +
  theme_minimal() 
```


## Secondary analysis- country-specific message effects

```{r}

res_ban_farming_subgroup = readRDS(here("results/res_farming_subgroup.rds"))
res_ban_sale_subgroup = readRDS(here("results/res_sale_subgroup.rds"))
res_ban_subsidies_subgroup = readRDS(here("results/res_subsidies_subgroup.rds"))

ggplot(res_ban_farming_subgroup, aes(x=Country, y=est, group=tr, color=tr)) + geom_point(position = position_dodge(0.5)) +
  geom_errorbar(aes(ymin=est.lb, ymax=est.ub), width=0.2, position = position_dodge(0.5)) +
  geom_hline(yintercept=0, linetype="dashed", color="red") +
  labs(title="Effect of Message on Support\nfor Ban on Farming Octopuses",
       x="Message Condition",
       y="Estimated Effect (95% CI)") +
  coord_flip() +
  theme_minimal() 

ggplot(res_ban_sale_subgroup, aes(x=Country, y=est, group=tr, color=tr)) + 
  geom_point(position = position_dodge(0.5)) +
  geom_errorbar(aes(ymin=est.lb, ymax=est.ub), width=0.2, position = position_dodge(0.5)) +
  geom_hline(yintercept=0, linetype="dashed", color="red") +
  labs(title="Effect of Message on Support\nfor Ban on Sale of Octopuses",
       x="Message Condition",
       y="Estimated Effect (95% CI)") +
  coord_flip() +
  theme_minimal() 

ggplot(res_ban_subsidies_subgroup, aes(x=Country, y=est, group=tr, color=tr)) + 
    geom_point(position = position_dodge(0.5)) +
  geom_errorbar(aes(ymin=est.lb, ymax=est.ub), width=0.2, position = position_dodge(0.5)) +
  geom_hline(yintercept=0, linetype="dashed", color="red") +
  labs(title="Effect of Message on Support\nfor Ban on Subsidies",
       x="Message Condition",
       y="Estimated Effect (95% CI)") +
  coord_flip() +
  theme_minimal() 



```


## Variable importance analysis- multivariate associations

```{r}
plotdf_Y1 = readRDS(here("results/res_farm_vim_glm.rds"))
plotdf_Y2 = readRDS(here("results/res_sale_vim_glm.rds"))
plotdf_Y3 = readRDS(here("results/res_subsidies_vim_glm.rds"))

ggplot(plotdf_Y1, aes(x=X, y=est )) + geom_point() +
  geom_errorbar(aes(ymin=ci.lb, ymax=ci.ub), width=0.5) +
  geom_hline(yintercept=0, linetype="dashed", color="red") +
  labs(title="Support for octopus farm ban",
    x="Covariate",
    y="Estimated Effect (95% CI)") +
  coord_flip() +theme_minimal() 

ggplot(plotdf_Y2, aes(x=X, y=est )) + geom_point() +
  geom_errorbar(aes(ymin=ci.lb, ymax=ci.ub), width=0.5) +
  geom_hline(yintercept=0, linetype="dashed", color="red") +
  labs(title="Support for octopus sale ban",
    x="Covariate",
    y="Estimated Effect (95% CI)") +
  coord_flip() +theme_minimal() 

ggplot(plotdf_Y3, aes(x=X, y=est )) + geom_point() +
  geom_errorbar(aes(ymin=ci.lb, ymax=ci.ub), width=0.5) +
  geom_hline(yintercept=0, linetype="dashed", color="red") +
  labs(title="Support for ban on subsidies",
    x="Covariate",
    y="Estimated Effect (95% CI)") +
  coord_flip() +theme_minimal() 

```

## Sensitivity results- Variable importance analysis- bivariate associations

```{r}
plotdf_Y1 = readRDS(here("results/res_farm_vim_bivariate.rds"))
plotdf_Y2 = readRDS(here("results/res_sale_vim_bivariate.rds"))
plotdf_Y3 = readRDS(here("results/res_subsidies_vim_bivariate.rds"))

ggplot(plotdf_Y1, aes(x=X, y=est )) + geom_point() +
  geom_errorbar(aes(ymin=ci.lb, ymax=ci.ub), width=0.5) +
  geom_hline(yintercept=0, linetype="dashed", color="red") +
  labs(title="Support for octopus farm ban",
    x="Covariate",
    y="Estimated Effect (95% CI)") +
  coord_flip() +theme_minimal() 

ggplot(plotdf_Y2, aes(x=X, y=est )) + geom_point() +
  geom_errorbar(aes(ymin=ci.lb, ymax=ci.ub), width=0.5) +
  geom_hline(yintercept=0, linetype="dashed", color="red") +
  labs(title="Support for octopus sale ban",
    x="Covariate",
    y="Estimated Effect (95% CI)") +
  coord_flip() +theme_minimal() 

ggplot(plotdf_Y3, aes(x=X, y=est )) + geom_point() +
  geom_errorbar(aes(ymin=ci.lb, ymax=ci.ub), width=0.5) +
  geom_hline(yintercept=0, linetype="dashed", color="red") +
  labs(title="Support for ban on subsidies",
    x="Covariate",
    y="Estimated Effect (95% CI)") +
  coord_flip() +theme_minimal() 

```

## Sensitivity results- unadjusted effects of message conditions 

```{r}

res_ban_farming_unadj = readRDS(here("results/res_farming_unadj.rds"))
res_ban_sale_unadj = readRDS(here("results/res_sale_unadj.rds"))
res_ban_subsidies_unadj = readRDS(here("results/res_subsidies_unadj.rds"))

ggplot(res_ban_farming_unadj, aes(x=Y, y=`Coef.`)) + geom_point() +
  geom_errorbar(aes(ymin=`2.5%`, ymax=`97.5%`), width=0.2) +
  geom_hline(yintercept=0, linetype="dashed", color="red") +
  labs(title="Effect of Message on Support\nfor Ban on Farming Octopuses",
       x="Message Condition",
       y="Estimated Effect (95% CI)") +
  coord_flip() +
  theme_minimal() 

ggplot(res_ban_sale_unadj, aes(x=Y, y=`Coef.`)) + geom_point() +
  geom_errorbar(aes(ymin=`2.5%`, ymax=`97.5%`), width=0.2) +
  geom_hline(yintercept=0, linetype="dashed", color="red") +
  labs(title="Effect of Message on Support\nfor Ban on Sale of Octopuses",
       x="Message Condition",
       y="Estimated Effect (95% CI)") +
  coord_flip() +
  theme_minimal() 

ggplot(res_ban_subsidies_unadj, aes(x=Y, y=`Coef.`)) + geom_point() +
  geom_errorbar(aes(ymin=`2.5%`, ymax=`97.5%`), width=0.2) +
  geom_hline(yintercept=0, linetype="dashed", color="red") +
  labs(title="Effect of Message on Support\nfor Ban on Subsidies",
       x="Message Condition",
       y="Estimated Effect (95% CI)") +
  coord_flip() +
  theme_minimal() 

```

